---
title: 'Reproducible Research: Peer Assessment 1'
author: "Muhammad Sadiq"
date: "June 11, 2016"
output: 
  html_document:
    keep_md: true
---
```{r}
library("dplyr")
library(lattice)
```

## Loading and preprocessing the data

```{r}
setwd("C:/Users/muhammad.sadiq/Documents")
data<-read.csv("activity.csv", na.strings = "NA")
head(data,2)
```
Checking the class of date
```{r}
class(data$date)
```
Converting date to date object
```{r}
data$date<-as.Date(data$date, format = "%Y-%m-%d")
class(data$date)
```
Checking the class of Interval
```{r}
class(data$interval)
```
Converting Interval class to factor
```{r}
data$interval<-factor(data$interval)
class(data$interval)
```

## What is mean total number of steps taken per day?
Remove "NA" from steps column
```{r}
cdata<-data[!is.na(data[,1]),]
head(cdata)

```
Following compute the mean number of steps on daily basis
```{r}
agg<-aggregate(steps~date, data = cdata, mean)
head(agg)
```

## What is the average daily activity pattern?

```{r}
hist(as.numeric(agg$steps), breaks = 20, col = "red", xlab = "Number of Steps", main= "Histogram of the total number of steps taken each day")
```

## Imputing missing values
It is assumed that "NA" for steps means no step so we replacing it with "0".
```{r}
idata<-data
idata[is.na(idata)]<-0
```

## Are there differences in activity patterns between weekdays and weekends?
Factoring Days
```{r}
idata$day <- as.factor(weekdays(idata$date))
str(idata$day)
```
Identifying Week day and weekends
```{r}
idata$week_day <- ifelse(!(idata$day %in% c("Saturday","Sunday")), TRUE, FALSE) 
```
Computing mean on weekdays
```{r}
weekdays_data <- idata[idata$week_day,]
steps_per_interval_weekdays <-aggregate(weekdays_data$steps,by=list(interval=weekdays_data$interval), FUN=mean)
```
Computing mean on week ends
```{r}
weekends_data <- idata[!idata$week_day,]
steps_per_interval_weekends <- aggregate(weekends_data$steps, by=list(interval=weekends_data$interval), FUN=mean)
```
Formatting and combining data
```{r}
colnames(steps_per_interval_weekdays) <- c("interval", "average_steps")
colnames(steps_per_interval_weekends) <- c("interval", "average_steps")
steps_per_interval_weekdays$day <- "Weekday"
steps_per_interval_weekends$day <- "Weekend"
week_data <- rbind(steps_per_interval_weekends, steps_per_interval_weekdays)
week_data$day <- as.factor(week_data$day)
```
Plotting the data
```{r}
xyplot(average_steps ~  interval | day, data = week_data, layout = c(1,2), type ="l", ylab="Number of Steps")
```
